Comparing Hydrogen and Jet-A for an N+3 Turbofan with Water Recirculation using Gradient-Based Optimization

### 2023-04-28 conference-paper/main.tex
Comments for the optimization problem and multipoint sections

\hypersetup{
    pdfauthor={Peter Atma}, % insert author here
	pdftitle={Aviation 2022 Abstract}, % insert title here
	pdfsubject={Propulsion Analysis and Optimization of an Ultra Bypass Turbofan Engine}, % insert keywords here
}

@@

% ==================================================
%	OVERALL COMMENTS
% ==================================================

@@

This work is organized as follows. First, in Section \ref{sec:method}, we introduce the turbofan model and explain the water injection and water recovery components.
In section \ref{sec:optprob} the implementation of the multipoint optimization problem is discussed.
Finally, we present the optimized results and discuss the design space in section \ref{sec:results}.

@@

\subsection{Multipoint Model}
We operate the engine cycle in different modes depending on the desired conservation relationships, design rules, and flight conditions.
In ``on-design'' mode, we prescribe important cycle inputs such as turbo machinery efficiencies, pressure ratios, and combustion temperatures.
The ``off-design'' mode inherits the flow path areas and turbo machinery map scalars from the ``on-design'' analysis.
We enforce design rules as balance equations between different design points to consider the performance of the engine cycle at limiting flight conditions.

% [ ] TODO: AL-PA, Add a small table to show the mach and altitude of each flight condition
The ``on-design'' point is top-of-climb (TOC) with ``off-design'' conditions at rolling takeoff (RTO), sea-level static (SLS), and cruise (CRZ).
Each flight condition is either critical to the performance or limits the design of the engine.
% [ ] TODO: AL-PA, You should explain why each design point is important in a separate sentence for each.
%   - For example, the SLS point is important for making sure the turbomachinery and flow paths are designed to meet a static thrust target
%   - The RTO point is limiting for temperature, etc...

% [ ] TODO: AL-PA, Consider re-writing this section using the comments from the previous TODO.  Be more concise and direct with the language here.
The N+3 engine model as it is implemented in pyCycle has many complex thermodynamic and performance connections between different operating conditions.
Therefore, the N+3 engine model computes the engine performance at 4 different operating conditions in order to resolve the complexities between each operating condition on the final design.
The operating conditions that considered are top-of-climb (TOC), rotating takeoff (RTO), sea-level static (SLS), and cruise (CRZ).
The SLS condition represents the static performance requirements for the engine, the RTO condition represents the performance and cooling requirements at the point of rotation during takeoff, the TOC operating condition represents the performance requirements during a climb, and the CRZ operation condition represents the criuse performance of the engine.
To account for each of these operating conditions, the N+3 model uses a technique called Multipoint Design Point (MDP) modeling to converge the model to an engine design that satisfies the requirements at each of these operating conditions.
One of these operating conditions is specified as the design point, which in this case is TOC, and the other operating conditions are considered the off-design points.
% NOTE: AL-PA, The design point sizes the flow paths and scales the turbomachinery.  The off-design has the exact same flow path as the design mode cycle.  The turbo machinery maps are scaled using scalar quantities for off-design performance.
The design point for the model generally sets all of the engine sizing such as areas and pressure ratios across components.
The off-design points are essentially the size of the engine at the design point but operating at the corresponding required thrust, cooling, and environmental conditions.
% [ ] TODO: AL-PA, We don't need to talk about the Newton or linear solver details.  Just explain the multipoint model structure with the Newton solver added for context.
% The design point of the engine model is converged using a linear Newton solver which then passes the geometric sizing variables such as cross-sectional areas and map scalars to the off-design points and similarly converges those.
% Then, a nonlinear Newton solver is used to converge the overall model with respect to the variable connections between each operating condition.
An XDSM diagram of the nominal N+3 engine model is shown in Figure \ref{fig:N3_xdsm} \cite{Hendricks2019}.

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.75\textwidth]{N3_xdsm.pdf}
    \caption{
        N+3 engine multipoint setup XDSM diagram.
        This XDSM diagram shows how each of the operating conditions is coupled with each other and how the outputs are computed.
    }
    \label{fig:N3_xdsm}
\end{figure}

@@

% Explain Exit ratios and how BPR and OD net thrusts are set
% [ ] TODO: AL-PA, consider removing this stuff and just direct the reader to the pyCycle paper.
In the N+3 model implementation presented in the pyCycle paper, the bypass ratio (BPR) at TOC is computed using a Balance component.
The Balance component for $BPR_{TOC}$ uses the exit velocity ratio of the core and bypass nozzles which is given as:

@@

\subsection{Performance Metrics}
Past implementations of engine cycles and example engine cycles in pyCycle generally have used either JetA or JP-7 as a fuel.
However, in this work we are interested in comparing the performance of similar engines with different types of fuel, specifically hydrogen due to the many benefits it offers in terms of performance and thermal properties.
Fortunately, pyCycle comes ready to use hydrogen with cataloged exhaust products for hydrogen fuel.
The fuel type is simply changed from Jet-A to H2 in the pyCycle model.

@@

\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.75\textwidth]{N3_xdsm_full.pdf}
    \caption{
        Full model N+3 XDSM diagram.
        This XDSM diagram shows the multipoint coupling between the different operation conditions and shows how the water recovery fractions are used to solve for water mass flow rates.
    }
    \label{fig:N3_xdsm_full}
\end{figure}

% Update with table from Andrew
\begin{table}[hbt!]
    \centering
    \caption{
        N+3 engine model parameters.
        Atmospheric parameters specify the flight conditions and humidity of the atmosphere and Lower Heating Values are given for each fuel to compute TSEC.}
    \begin{tabular}{c c c c}
        \hline
        Parameter              & Value   & Units          & Comments                                        \\
        \hline
        $h_{\rm{TOC}}$         & 0.00017 & -              & humidity ratio at TOC \cite{Kalnay1996}         \\
        $h_{\rm{CRZ}}$         & 0.00017 & -              & humidity ratio at CRZ \cite{Kalnay1996}         \\
        $h_{\rm{RTO}}$         & 0.009   & -              & humidity ratio at RTO \cite{Kalnay1996}         \\
        $h_{\rm{SLS}}$         & 0.009   & -              & humidity ratio at SLS \cite{Kalnay1996}         \\
        $\rm{LHV}_{\rm{JetA}}$ & 18564.0 & \unit{BTU/lbm} & Lower heating value of Jet-A \cite{Lawicki2002} \\
        $\rm{LHV}_{\rm{H2}}$   & 51591.0 & \unit{BTU/lbm} & Lower heating value of H2 \cite{toolbox2003}    \\
        \hline
    \end{tabular}
    \label{engine_params}
\end{table}

@@

\subsection{Problem Statment}
% Briefly discuss the optimization problem
We will perform gradient-based design optimization of the engine model with the multipoint architecture that considers each of the flight conditions experienced by commercial aircraft.
% We will perform a parameter sweep of the limiting constraint to understand the impact of different emissions and performance requirements on the design space.
The objective is to minimize the TSEC since that is our performance metric for comparing the two types of fuels.
However, since we are constraining net thrust at each operating condition this optimization problem becomes ill-posed.
Therefore, we will minimize fuel flow rate, $W_{fuel}$ at the cruise condition subject to net thrust and engine diameter constraints since this will accomplish the same goal with a better posed problem.
An XDSM diagram of the optimization problem with the multipoint formulation is shown in Figure \ref{fig:N3_xdsm_opt}.

\begin{figure}[!hbt]
    \centering
    \includegraphics[width=0.75\textwidth]{N3_xdsm_opt.pdf}
    \caption{
        XDSM diagram of the multipoint optimization problem with constraints.
        The XDSM diagram shows the variables and outputs within the optimization model and how each of these values is connected to the optimizer and design points.}
    \label{fig:N3_xdsm_opt}
\end{figure}

This optimization problem was run for both Jet-A and H2 fuels using the parameters given in Table \ref{engine_params}.
The optimization problem objective function, design variables, and constraints are shown in Table \ref{tab:opt_problem} below.

% [ ] TODO: AL - PA, Define the T4 ratio with an equation and reference the equation in the table
% [ ] TODO: AL-PA, I thought we were only doing recovery fractions at CRZ??
\begin{table}[hbt!]
    \centering
    \caption{
        Multipoint optimization problem definition.
        The 4 design variables are the water recovery fractions of the core exhaust stream at each operation condition.
        The objective function is the thrust specific energy consumption at the cruise condition with a net thrust constraint at the design point, TOC.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l}
        \toprule
                        & Variable/Function              & Description                                 & Units            & Quantity \\
        \hline
        minimize        & $ \dot{m}_{\rm{fuel,CRZ}} $    & Fuel flow rate at CRZ                       & \unit{lbm/s}     & 1        \\
                        &                                &                                             &                  &          \\
        with respect to & $X_{\rm{H2O,TOC}}$             & Water recovery fraction at TOC              & -                & 1        \\
                        & $X_{\rm{H2O,RTO}}$             & Water recovery fraction at RTO              & -                & 1        \\
                        & $X_{\rm{H2O,SLS}}$             & Water recovery fraction at SLS              & -                & 1        \\
                        & $X_{\rm{H2O,CRZ}}$             & Water recovery fraction at CRZ              & -                & 1        \\
                        & $\rm{PR}_{\rm{HPC,TOC}}$       & TOC high-pressure compressor pressure ratio & -                & 1        \\
                        & $\rm{PR}_{\rm{LPC,TOC}}$       & TOC low-pressure compressor pressure ratio  & -                & 1        \\
                        & $\rm{PR}_{\rm{fan,TOC}}$       & TOC fan pressure ratio                      & -                & 1        \\
                        & $T_{4,\rm{ratio}}$             & TOC-to-RTO temperature ratio                & -                & 1        \\
                        & $T_{4,\rm{RTO}}$               & RTO combustor temperature                   & \unit{\degree R} & 1        \\
                        & $V_{\rm{ratio,CRZ}}$           & RTO combustor temperature                   & \unit{\degree R} & 1        \\
        \cline{3-5}
                        &                                & Total                                       &                  & 10       \\
                        &                                &                                             &                  &          \\
        subject to      & $F_{\rm{net,TOC}} \geq 5800.0$ & Target net thrust at TOC                    & \unit{lbf}       & 1        \\
                        & $D_{\rm{fan}} \leq 100$        & Maximum Fan Diameter                        & \unit{inch^2}    & 1        \\
        \cline{3-5}
                        &                                & Total                                       &                  & 2        \\
        \bottomrule
    \end{tabular}
    \label{tab:opt_problem}
\end{table}

% Blurb about PyOptSparse and SNOPT
\subsection{Optimization Software}
OpenMDAO supports the use of different optimization packages including Scipy and PyOptSparse.
PyOptSparse is an object-oriented framework for formulating and solving nonlinear constrained optimization problems.
PyOptSparse was chosen as the main optimization package for this problem due to its wide range of open-source gradient-based optimizers and integration with  \cite{Wu2020a}.

SNOPT (Sparse Nonlinear OPTimizer) was the optimizer chosen for this work.
SNOPT is a state-of-the-art optimizer used to solve large-scale optimization problems (linear and nonlinear).
The software uses a sparse SQP (sequential quadratic programming) algorithm with a limited-memory quasi-Newton approximation to the Hessian of the lagrangian.
SNOPT is particularly useful for nonlinear problems whose functions and derivatives are expensive compute.
An augmented Lagrangian merit function ensures convergence from an arbitrary point.
SNOPT allows the nonlinear constraints to be violated (if necessary) and minimizes the sum of these violations.
Due to the robustness of the algorithm, SNOPT is a perfect optimizer to solve the highly complex Multipoint systek since the pyCycle engine cycle would occasionally reach a point that is non-physical and would thus cause the solver to fail.
SNOPT is able to respond to these solver failures and backtrack to a region in the design space that is physical and continue the optimization \cite{Gill2005a}.

>>>>

\hypersetup{
    pdfauthor={Peter Atma}, % insert author here
	pdftitle={Aviation 2023 Paper}, % insert title here
	pdfsubject={Comparing Hydrogen and Jet-A for an N+3 Turbofan with Water Recirculation using Gradient-Based Optimization}, % insert keywords here
}

@@

% ==================================================
%	OVERALL TODOS
% ==================================================
% [ ] TODO: AL-PA, Need to add equation labels
% [ ] TODO: AL-PA, When you are citing things you should include a "~" to link the citation to the text.  i.e, "OpenMDAO~\cite{Gray2019a}" (See my edits in the introduction)
% [ ] TODO: AL-PA, You should also use the "~" to link all equation, figure, and table references to the text.
% [ ] TODO: AL-PA, Make sure all acronyms and subscripts that are longer than a single letter or represent words are shown as text using \rm{} in math mode

@@

This work is organized as follows. First, in Section~\ref{sec:method}, we introduce the turbofan model and explain the water injection and water recovery components.
In section~\ref{sec:optprob} the implementation of the multipoint optimization problem is discussed.
Finally, we present the optimized results and discuss the design space in section~\ref{sec:results}.

@@

\subsection{Multipoint Model}
% NOTE: AL-PA, I added this blurb to introduce multipoint
We operate the engine cycle in different modes depending on the desired conservation relationships, design rules, and flight conditions.
In ``on-design'' mode, we prescribe important cycle inputs such as turbo machinery efficiencies, pressure ratios, and combustion temperatures.
The ``off-design'' mode inherits the flow path areas and turbo machinery map scalars from the ``on-design'' analysis.
We enforce design rules as balance equations between different design points to consider the performance of the engine cycle at limiting flight conditions.

% [ ] TODO: AL-PA, Add a small table to show the mach and altitude of each flight condition
The ``on-design'' point is top-of-climb (TOC) with ``off-design'' conditions at rolling takeoff (RTO), sea-level static (SLS), and cruise (CRZ).
Each flight condition is either critical to the performance or limits the design of the engine.
% [ ] TODO: AL-PA, You should explain why each design point is important in a separate sentence for each.
%   - For example, the SLS point is important for making sure the turbo machinery and flow paths are designed to meet a static thrust target
%   - The RTO point is limiting for temperature, etc...

% [ ] TODO: AL-PA, Consider re-writing this section using the comments from the previous TODO.  Be more concise and direct with the language here.
The N+3 engine model as it is implemented in pyCycle has many complex thermodynamic and performance connections between different operating conditions.
Therefore, the N+3 engine model computes the engine performance at 4 different operating conditions in order to resolve the complexities between each operating condition on the final design.
The operating conditions that considered are top-of-climb (TOC), rotating takeoff (RTO), sea-level static (SLS), and cruise (CRZ).
% [ ] TODO: AL-PA, This is a big run-on sentence.  Consider breaking it up and being more specific about the requirements at each design point.  See previous comments.
The SLS condition represents the static performance requirements for the engine, the RTO condition represents the performance and cooling requirements at the point of rotation during takeoff, the TOC operating condition represents the performance requirements during a climb, and the CRZ operation condition represents the criuse performance of the engine.
To account for each of these operating conditions, the N+3 model uses a technique called Multipoint Design Point (MDP) modeling to converge the model to an engine design that satisfies the requirements at each of these operating conditions.
% NOTE: AL-PA, Using a lot of extra words to say: "The TOC point is the on-design condition."
One of these operating conditions is specified as the design point, which in this case is TOC, and the other operating conditions are considered the off-design points.
% NOTE: AL-PA, The design point sizes the flow paths and scales the turbo machinery.  The off-design has the exact same flow path as the design mode cycle.  The turbo machinery maps are scaled using scalar quantities for off-design performance.
% TODO: AL-PA, Don't say "generally" or "essentially" because it makes the method appear ambiguous.
The design point for the model generally sets all of the engine sizing such as areas and pressure ratios across components.
The off-design points are essentially the size of the engine at the design point but operating at the corresponding required thrust, cooling, and environmental conditions.
% [ ] TODO: AL-PA, We don't need to talk about the Newton or linear solver details.  Just explain the multipoint model structure with the Newton solver added for context.
% The design point of the engine model is converged using a linear Newton solver which then passes the geometric sizing variables such as cross-sectional areas and map scalars to the off-design points and similarly converges those.
% Then, a nonlinear Newton solver is used to converge the overall model with respect to the variable connections between each operating condition.
An XDSM diagram of the nominal N+3 engine model is shown in Figure \ref{fig:N3_xdsm} \cite{Hendricks2019}.

% [ ] TODO: AL-PA, I would remove this figure.  Not adding anything more than the detailed XDSM you provide later.
\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.75\textwidth]{N3_xdsm.pdf}
    \caption{
        N+3 engine multipoint setup XDSM diagram.
        This XDSM diagram shows how each of the operating conditions is coupled with each other and how the outputs are computed.
    }
    \label{fig:N3_xdsm}
\end{figure}

@@

% Explain Exit ratios and how BPR and OD net thrusts are set
% [ ] TODO: AL-PA, I would include all four of the balance components here and explain how they effectively act as constraints on the multipoint model.
% NOTE: AL-PA, See the pyCycle paper, they explain how using the balance residuals creates a reduced design space for the optimization problem.
% NOTE: AL-PA, You can discuss how this makes the multipoint nonlinear solution more challenging, but improves the optimization problem conditioning.
In the N+3 model implementation presented in the pyCycle paper, the bypass ratio (BPR) at TOC is computed using a Balance component.
The Balance component for $BPR_{TOC}$ uses the exit velocity ratio of the core and bypass nozzles which is given as:

@@

% [ ] TODO: AL-PA, I'm not sure this subsection is necessary.  We will describe the performance metrics in the optimization problem section.  Consider moving it there.
\subsection{Performance Metrics}
Past implementations of engine cycles and example engine cycles in pyCycle generally have used either JetA or JP-7 as a fuel.
However, in this work we are interested in comparing the performance of similar engines with different types of fuel, specifically hydrogen due to the many benefits it offers in terms of performance and thermal properties.
Fortunately, pyCycle comes ready to use hydrogen with cataloged exhaust products for hydrogen fuel.
The fuel type is simply changed from Jet-A to H2 in the pyCycle model.

@@

% [ ] TODO: AL-PA, Already discussed, but this should be the only "fine-detail" XDSM in the paper.
% [ ] TODO: AL-PA, Make the figure use the entire text width so that it's easier to read.  Might have to reposition a bit.
\begin{figure}[hbt!]
    \centering
    \includegraphics[width=0.75\textwidth]{N3_xdsm_full.pdf}
    \caption{
        Full model N+3 XDSM diagram.
        This XDSM diagram shows the multipoint coupling between the different operation conditions and shows how the water recovery fractions are used to solve for water mass flow rates.
    }
    \label{fig:N3_xdsm_full}
\end{figure}

% Update with table from Andrew
% [ ] TODO: AL-PA, Why are some of the fist words in the descriptions capitalized but not others?  I think this should be consistent.
% [ ] TODO: AL-PA, Change the lower heating value references to footnotes. i.e \footnote{\url{...}}
\begin{table}[hbt!]
    \centering
    \caption{
        N+3 engine model parameters.
        Atmospheric parameters specify the flight conditions and humidity of the atmosphere and Lower Heating Values are given for each fuel to compute TSEC.}
    \begin{tabular}{r r r l}
        \hline
        Parameter              & Value   & Units          & Description                                     \\
        \hline
        $h_{\rm{TOC}}$         & 0.00017 & -              & humidity ratio at TOC \cite{Kalnay1996}         \\
        $h_{\rm{CRZ}}$         & 0.00017 & -              & humidity ratio at CRZ \cite{Kalnay1996}         \\
        $h_{\rm{RTO}}$         & 0.009   & -              & humidity ratio at RTO \cite{Kalnay1996}         \\
        $h_{\rm{SLS}}$         & 0.009   & -              & humidity ratio at SLS \cite{Kalnay1996}         \\
        $\rm{LHV}_{\rm{JetA}}$ & 18564.0 & \unit{BTU/lbm} & Lower heating value of Jet-A \cite{Lawicki2002} \\
        $\rm{LHV}_{\rm{H2}}$   & 51591.0 & \unit{BTU/lbm} & Lower heating value of H2 \cite{toolbox2003}    \\
        \hline
    \end{tabular}
    \label{engine_params}
\end{table}

@@

% NOTE: AL-PA, No reason to have an extra subsection here so I removed it
% \subsection{Problem Statment}
% Briefly discuss the optimization problem
% NOTE: AL-PA, "We will perform" makes it seem like we haven't done the work yet.
We performed multipoint gradient-based design optimization of the N+3 engine model considering TOC, RTO, SLS, and CRZ flight conditions.
% We will perform gradient-based design optimization of the engine model with the multipoint architecture that considers each of the flight conditions experienced by commercial aircraft.
% We will perform a parameter sweep of the limiting constraint to understand the impact of different emissions and performance requirements on the design space.
% NOTE: AL-PA, I re-wrote this to be more direct
The objective is to minimize fuel burn subject to design and performance constraints at each flight condition.
% The objective is to minimize the TSEC since that is our performance metric for comparing the two types of fuels.
% However, since we are constraining net thrust at each operating condition this optimization problem becomes ill-posed.
% Therefore, we will minimize fuel flow rate, $W_{fuel}$ at the cruise condition subject to net thrust and engine diameter constraints since this will accomplish the same goal with a better posed problem.
% [ ] TODO: AL-PA, This is a good place to discuss the performance metrics and add the equations for TSEC and explain why it's better for this study than TSFC.
An XDSM diagram of the optimization problem with the multipoint formulation is shown in Figure \ref{fig:N3_xdsm_opt}.

% [ ] TODO: AL-PA, simplify this XDSM to show just the optimization setup. Condense the MDA into a single block and remove all the multipoint acronyms.
\begin{figure}[!hbt]
    \centering
    \includegraphics[width=0.75\textwidth]{N3_xdsm_opt.pdf}
    \caption{
        XDSM diagram of the multipoint optimization problem with constraints.
        The XDSM diagram shows the variables and outputs within the optimization model and how each of these values is connected to the optimizer and design points.}
    \label{fig:N3_xdsm_opt}
\end{figure}

% [ ] TODO: AL-PA, I would elaborate on this a bit more.  Are these the on-design parameters?  Are they constraints?
% [ ] TODO: AL-PA, The N+3 model has a bunch of constraints imposed on the multipoint problem via balance comps.  This is a good place to talk about those.
% [ ] TODO: AL-PA, Discuss how using balance comps instead of constraints represents a reduced space problem for the optimizer. (See pycycle paper for more information).
This optimization problem was run for both Jet-A and H2 fuels using the parameters given in Table \ref{engine_params}.
The optimization problem objective function, design variables, and constraints are shown in Table \ref{tab:opt_problem} below.

% [ ] TODO: AL - PA, Define the T4 ratio with an equation and reference the equation in the table
% [ ] TODO: AL-PA, I thought we were only doing recovery fractions at CRZ??
% [ ] TODO: AL-PA, Make sure you reference the equations for the DVs/cons in the descriptions so the reader can connect them.
\begin{table}[hbt!]
    \centering
    \caption{
        Multipoint optimization problem definition.
        The 4 design variables are the water recovery fractions of the core exhaust stream at each operation condition.
        The objective function is the thrust specific energy consumption at the cruise condition with a net thrust constraint at the design point, TOC.
    }
    \small
    \renewcommand{\arraystretch}{1.2}
    \begin{tabular}{r l l l l}
        \toprule
                        & Variable/Function              & Description                                 & Units            & Quantity \\
        \hline
        minimize        & $ \dot{m}_{\rm{fuel,CRZ}} $    & Fuel flow rate at CRZ                       & \unit{lbm/s}     & 1        \\
                        &                                &                                             &                  &          \\
        with respect to & $X_{\rm{H2O,TOC}}$             & Water recovery fraction at TOC              & -                & 1        \\
                        & $X_{\rm{H2O,RTO}}$             & Water recovery fraction at RTO              & -                & 1        \\
                        & $X_{\rm{H2O,SLS}}$             & Water recovery fraction at SLS              & -                & 1        \\
                        & $X_{\rm{H2O,CRZ}}$             & Water recovery fraction at CRZ              & -                & 1        \\
                        & $\rm{PR}_{\rm{HPC,TOC}}$       & TOC high-pressure compressor pressure ratio & -                & 1        \\
                        & $\rm{PR}_{\rm{LPC,TOC}}$       & TOC low-pressure compressor pressure ratio  & -                & 1        \\
                        & $\rm{PR}_{\rm{fan,TOC}}$       & TOC fan pressure ratio                      & -                & 1        \\
                        & $T_{4,\rm{ratio}}$             & TOC-to-RTO temperature ratio                & -                & 1        \\
                        & $T_{4,\rm{RTO}}$               & RTO combustor temperature                   & \unit{\degree R} & 1        \\
                        & $V_{\rm{ratio,CRZ}}$           & Core-to-bypass nozzle velocity ratio at CRZ & \unit{\degree R} & 1        \\
        \cline{3-5}
                        &                                & Total                                       &                  & 10       \\
                        &                                &                                             &                  &          \\
        subject to      & $F_{\rm{net,TOC}} \geq 5800.0$ & Target net thrust at TOC                    & \unit{lbf}       & 1        \\
                        & $D_{\rm{fan}} \leq 100$        & Maximum Fan Diameter                        & \unit{inch^2}    & 1        \\
        \cline{3-5}
                        &                                & Total                                       &                  & 2        \\
        \bottomrule
    \end{tabular}
    \label{tab:opt_problem}
\end{table}

% Blurb about PyOptSparse and SNOPT
% NOTE: AL-PA, This can all be part of the optimization problem section you don't need a new subsection
% \subsection{Optimization Software}
% NOTE: AL-PA, "pyOptSparse" not "PyOptSparse", i.e the first "p" is not capitalized.
% [x] TODO: AL-AL, need to cite pyOptSparse, SNOPT, and OpenMDAO
% NOTE: AL-PA, I re-wrote most of this to be more concise.
OpenMDAO~\cite{Gray2019a} is a modular framework that enables multidisciplinary gradient-based optimization with analytic coupled derivatives.
We use pyOptSparse~\cite{Wu2020a} to facilitate the use of state-of-the-art optimization software through a unified python interface.
We solve the optimization problem listed in Table~\ref{tab:opt_problem} with SNOPT~\cite{Gill2005a}, a gradient-based sequential quadratic programming (SQP) algorithm for large-scale constrained problems.

% OpenMDAO supports the use of different optimization packages including Scipy and PyOptSparse.
% PyOptSparse is an object-oriented framework for formulating and solving nonlinear constrained optimization problems.
% PyOptSparse was chosen as the main optimization package for this problem due to its wide range of open-source gradient-based optimizers and integration with  \cite{Wu2020a}.

% SNOPT (Sparse Nonlinear OPTimizer) was the optimizer chosen for this work.
% SNOPT is a state-of-the-art optimizer used to solve large-scale optimization problems (linear and nonlinear).
% The software uses a sparse SQP (sequential quadratic programming) algorithm with a limited-memory quasi-Newton approximation to the Hessian of the lagrangian.
% SNOPT is particularly useful for nonlinear problems whose functions and derivatives are expensive compute.
% An augmented Lagrangian merit function ensures convergence from an arbitrary point.
% SNOPT allows the nonlinear constraints to be violated (if necessary) and minimizes the sum of these violations.
% Due to the robustness of the algorithm, SNOPT is a perfect optimizer to solve the highly complex Multipoint systek since the pyCycle engine cycle would occasionally reach a point that is non-physical and would thus cause the solver to fail.
% SNOPT is able to respond to these solver failures and backtrack to a region in the design space that is physical and continue the optimization \cite{Gill2005a}.
